---
title: "SA.Project"
author: "Molly McCann"
date: "2024-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(StatsBombR)
library(dplyr)
library(SBpitch)
library(grid)
library(ineq)
library(Gini)
library(ggplot2) 
library(ggdark) 
library(ggsoccer) 
library(xgboost) 
library(fastDummies) 
library(xgboost) 
library(reshape) 
library(purrr)
library(reshape2)
```

```{r}
comp <- FreeCompetitions() |>
filter(competition_id==37 & season_name=="2019/2020")

matches <- FreeMatches(comp)

statsBombData <- free_allevents(MatchesDF = matches, Parallel = T)

statsBombData = allclean(statsBombData)
```

```{r}
unique(statsBombData$shot.technique.name)
unique(statsBombData$pass.technique.name)
unique(statsBombData$team.name)
unique(statsBombData$tactics.formation)
```

## Match IDs for Chelsea FCW
```{r}
chelseaMatches <- matches |>
  filter(home_team.home_team_name == "Chelsea FCW" | away_team.away_team_name == "Chelsea FCW")

unique(chelseaMatches$match_id)
```

## Selecting Chelsea FCW match data 
```{r}
chelseaAttackData <- statsBombData |>
  filter(match_id %in% c(2275154, 2275105, 2275063, 2275072, 2275057, 
                         2275045, 2275093, 2275084, 2275052, 2275080, 
                         2275090, 2275108, 2275132, 2275130, 2275074))
```

## Add column with stadium.id to chelseaAttackData
```{r}
# Ensure chelseaAttackData has a column named 'stadium.id' by joining with chelseaMatches on 'match_id'
chelseaAttackData <- chelseaAttackData |>
  left_join(chelseaMatches |> 
              select(match_id, stadium.id), by = "match_id")

```

## Player names for Chelsea FCW
```{r}
chelseaPlayerData <- chelseaAttackData |>
  filter(team.name == "Chelsea FCW")

chelseaNames <- na.omit(unique(chelseaPlayerData$player.name))
print(chelseaNames)
```

## Millie Bright's defensive passes from the box

Boiler plate code produced by StatsBomb

```{r}
# Finding the player.id associated with Millie Bright (defender)
brightID <- chelseaAttackData |>
  filter(player.name == "Millie Bright") |>
  pull(player.id)

# Creating the visualization
brightPass <- chelseaAttackData |>
  filter(type.name == "Pass" & is.na(pass.outcome.name) & player.id == brightID) |>
  filter(location.x <= 18 & location.y <= 62 & location.y >= 18) |>
  filter(pass.end_location.x > 18 | pass.end_location.y > 62 | pass.end_location.y < 18)

create_Pitch(grass_colour = "#538032", 
             line_colour =  "#ffffff", 
             background_colour = "#538032", 
             goal_colour = "#000000") + 
  geom_segment(data = brightPass, aes(x = location.x, y = location.y,
                                    xend = pass.end_location.x, yend = pass.end_location.y), 
               lineend = "round", size = 0.5, color = "#ffffff", 
               arrow = arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Millie Bright, Completed Box Clearance/Pass", subtitle = "WSL, 2019-20") + 
  scale_y_reverse() + 
  coord_fixed(ratio = 105/100)
```

As a right center back, Millie Bright frequently passes to her fullbacks or wingers on the right side of the field. Her passes are very routine and systematic, which is a good trait of a dependable center back. Unlike attacking players who rely on creativity when the ball is at their feet, Millie is able to make safe and quality passes to advance Chelsea's play up the field. 

## Sam Kerr's attacking passes into the box 

Boiler plate code produced by StatsBomb

```{r}
# Finding the player.id associated with Sam Kerr (forward)
kerrID <- chelseaAttackData |>
  filter(player.name == "Samantha May Kerr") |>
  pull(player.id)

# Creating the visualization
kerrPass <- chelseaAttackData |>
  filter(type.name == "Pass" & is.na(pass.outcome.name) & player.id == kerrID) |>
  filter(pass.end_location.x >= 102 & pass.end_location.y <= 62 & pass.end_location.y >= 18)


create_Pitch(grass_colour = "#538032", 
             line_colour =  "#ffffff", 
             background_colour = "#538032", 
             goal_colour = "#000000") + 
  geom_segment(data = kerrPass, aes(x = location.x, y = location.y,
                                    xend = pass.end_location.x, yend = pass.end_location.y), 
               lineend = "round", size = 0.5, color = "#ffffff", 
               arrow = arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Sam Kerr, Completed Box Passes", subtitle = "WSL, 2019-20") + 
  scale_y_reverse() + 
  coord_fixed(ratio = 105/100)
```

This visualization shows the creative nature of Sam Kerr when she is attacking. Rather than a systematic series of passes seen in defenders, Sam Kerr's passes into the box vary and come from both the right and left sides of the field. This shows that she is comfortable with the movement neccessary for forwards to have when attacking. 

## Bethany England's attacking passes into the box 

Boiler plate code produced by StatsBomb

```{r}
# Finding the player.id associated with Sam Kerr (forward)
englandID <- chelseaAttackData |>
  filter(player.name == "Bethany England") |>
  pull(player.id)

# Creating the visualization
englandPass <- chelseaAttackData |>
  filter(type.name == "Pass" & is.na(pass.outcome.name) & player.id == englandID) |>
  filter(pass.end_location.x >= 102 & pass.end_location.y <= 62 & pass.end_location.y >= 18)


create_Pitch(grass_colour = "#538032", 
             line_colour =  "#ffffff", 
             background_colour = "#538032", 
             goal_colour = "#000000") + 
  geom_segment(data = englandPass, aes(x = location.x, y = location.y,
                                    xend = pass.end_location.x, yend = pass.end_location.y), 
               lineend = "round", size = 0.5, color = "#ffffff", 
               arrow = arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Bethany England, Completed Box Passes", subtitle = "WSL, 2019-20") + 
  scale_y_reverse() + 
  coord_fixed(ratio = 105/100)
```

This visualization displays the routes for Beth England's passes. As seen on the left side of the goal, Beth England is comfortable with passing the ball backwards at an angle from the endline in order to give her teammates a better angle to the goal. She also tends to dish the ball out to teammates around the penalty area, giving them an clear view of the goal. 

## So-Yun Ji's attacking passes into the box 

Boiler plate code produced by StatsBomb

```{r}
# Finding the player.id associated with Sam Kerr (forward)
jiID <- chelseaAttackData |>
  filter(player.name == "So-Yun Ji") |>
  pull(player.id)

# Creating the visualization
jiPass <- chelseaAttackData |>
  filter(type.name == "Pass" & is.na(pass.outcome.name) & player.id == jiID) |>
  filter(pass.end_location.x >= 102 & pass.end_location.y <= 62 & pass.end_location.y >= 18)


create_Pitch(grass_colour = "#538032", 
             line_colour =  "#ffffff", 
             background_colour = "#538032", 
             goal_colour = "#000000") + 
  geom_segment(data = jiPass, aes(x = location.x, y = location.y,
                                    xend = pass.end_location.x, yend = pass.end_location.y), 
               lineend = "round", size = 0.5, color = "#ffffff", 
               arrow = arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "So-Yun Ji, Completed Box Passes", subtitle = "WSL, 2019-20") + 
  scale_y_reverse() + 
  coord_fixed(ratio = 105/100)
```

## Guro Reiten's attacking passes into the box 

Boiler plate code produced by StatsBomb

```{r}
# Finding the player.id associated with Sam Kerr (forward)
reitenID <- chelseaAttackData |>
  filter(player.name == "Guro Reiten") |>
  pull(player.id)

# Creating the visualization
reitenPass <- chelseaAttackData |>
  filter(type.name == "Pass" & is.na(pass.outcome.name) & player.id == reitenID) |>
  filter(pass.end_location.x >= 102 & pass.end_location.y <= 62 & pass.end_location.y >= 18)


create_Pitch(grass_colour = "#538032", 
             line_colour =  "#ffffff", 
             background_colour = "#538032", 
             goal_colour = "#000000") + 
  geom_segment(data = reitenPass, aes(x = location.x, y = location.y,
                                    xend = pass.end_location.x, yend = pass.end_location.y), 
               lineend = "round", size = 0.5, color = "#ffffff", 
               arrow = arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Guro Reiten, Completed Box Passes", subtitle = "WSL, 2019-20") + 
  scale_y_reverse() + 
  coord_fixed(ratio = 105/100)
``` 


## Heatmap of defensive positioning for each team

Boiler plate code produced by StatsBomb

```{r}
# Get rid of values outside of the pitch dimensions 
heatmap <- statsBombData |>
  mutate(location.x = ifelse(location.x > 120, 120, location.x), 
         location.y = ifelse(location.y > 80, 80, location.y),
         location.x = ifelse(location.x < 0, 0, location.x),
         location.y = ifelse(location.y < 0, 0, location.y)) 

# Create the buckets needed to plot the zones
heatmap$xbin <- cut(heatmap$location.x, breaks = seq(from = 0, to = 120, by = 20), include.lowest = TRUE) 
heatmap$ybin <- cut(heatmap$location.y, breaks = seq(from = 0, to = 80, by = 20), include.lowest = TRUE) 

# 
heatmap <- heatmap |>
  filter(type.name == "Pressure" | duel.type.name == "Tackle" | type.name == "Foul Committed" |
           type.name == "Interception" | type.name == "Block" ) |>
  group_by(team.name) |>
  mutate(total_DA = n()) |>
  group_by(team.name, xbin, ybin) |>
  summarise(total_DA = max(total_DA),
            bin_DA = n(),
            bin_pct = bin_DA/total_DA,
            location.x = median(location.x),
            location.y = median(location.y)) |>
  group_by(xbin, ybin) |>
  mutate(league_ave = mean(bin_pct)) |>
  group_by(team.name, xbin, ybin) |>
  mutate(diff_vs_ave = bin_pct - league_ave) 

defensiveactivitycolors <- c("#dc2429", "#dc2329", "#df272d", "#df3238", "#e14348", "#e44d51", "#e35256", "#e76266", "#e9777b", "#ec8589",
                             "#ec898d", "#ef9195","#ef9ea1", "#f0a6a9", "#f2abae", "#f4b9bc", "#f8d1d2", "#f9e0e2","#f7e1e3", "#f5e2e4",
                             "#d4d5d8", "#d1d3d8", "#cdd2d6", "#c8cdd3", "#c0c7cd","#b9c0c8", "#b5bcc3", "#909ba5", "#8f9aa5", "#818c98",
                             "#798590", "#697785", "#526173", "#435367", "#3a4b60", "#2e4257", "#1d3048", "#11263e", "#11273e", "#0d233a","#020c16") 
ggplot(data = heatmap, aes(x = location.x, y = location.y, fill = diff_vs_ave, group =diff_vs_ave)) +
  geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) +
  annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, color = "black", size = 0.6) + 
  annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, color = "black", size = 0.6) + 
  annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, color = "white", size = 0.6) + 
  annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, color = "white", size = 0.6) + 
  annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, color = "white", size = 0.6) + 
  annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, color = "white", size = 0.6) + 
  annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, color = "black", size = 0.6) + 
  annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, color = "black", size = 0.6) + 
  annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, color = "white", size = 0.6) + 
  annotate("segment", x = 0, xend = 0, y = 0, yend = 80, color = "black", size = 0.6) + 
  annotate("segment", x = 120, xend = 120, y = 0, yend = 80, color = "black", size = 0.6) +
  theme(rect = element_blank(), line = element_blank()) +
  annotate("point", x = 12 , y = 40, color = "white", size = 1.05) + 
  annotate("point", x = 108 , y = 40, color = "white", size = 1.05) +
  annotate("path", color = "white", size = 0.6, x=60+10*cos(seq(0,2*pi,length.out=2000)), y=40+10*sin(seq(0,2*pi,length.out=2000)))+
  annotate("point", x = 60 , y = 40, color = "white", size = 1.05) + 
  annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6, y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") + 
  annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6, y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
  theme(axis.text.x=element_blank(), axis.title.x = element_blank(),axis.title.y = element_blank(),
        plot.caption=element_text(size=8,family="Arial", hjust=0.5, vjust=0.5), 
        plot.subtitle = element_text(size = 8, family="Arial", hjust = 0.5), 
        axis.text.y=element_blank(), legend.title = element_blank(), legend.text=element_text(size=8,family="Arial"), 
        legend.key.size = unit(1.5, "cm"), 
        axis.ticks=element_blank(), plot.background = element_rect(fill = "white"), 
        strip.text.x = element_text(size=8,family="Arial")) +
  scale_y_reverse() + 
  scale_fill_gradientn(colours = defensiveactivitycolors, trans = "reverse", labels = scales::percent_format(accuracy = 1), 
                       limits = c(0.03, -0.03)) +
  labs(title = "Where Do Teams Defend vs League Average?", subtitle = "FA Women's Super League, 2019/20") + 
  coord_fixed(ratio = 95/100) + 
  annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last", length=unit(2.55,"mm")), 
                                     gp=gpar(col="black", fill=NA, lwd=2.2)),
                    xmin=25, xmax = 95, ymin = -83, ymax = -83) + 
  facet_wrap(~team.name) + 
  guides(fill = guide_legend(reverse = TRUE))

```

The darker red zones indicate a higher concentration of defensive actions compared to the league average in those areas. For top teir teams like Chelsea and Arsenal, their defensive half has less defensive actions while their attacking half has a greater concentration of defensive actions when compared to the league average. This indicates that these teams attempt to win the ball back on their opponents defensive half. They are high pressure teams and tend to not let the opponent enter their defensive half frequently; thus, these zones have less defensive actions than the league average. 

## Chelsea Expected Goals [xG] by Position Group 

```{r}
# Separate data for shots only 
chelseaShotsData <- chelseaAttackData |>
  filter(type.name == "Shot" & team.name == "Chelsea FCW")
```

```{r}
# Binary operator for goals vs no goal 
chelseaShotsData$shot_res <- 0
chelseaShotsData$shot_res[chelseaShotsData$shot.outcome.name == "Goal"] <- 1
```

```{r}
# Look at different shot outcomes 
summary(as.factor(chelseaShotsData$shot.outcome.name))

# Look at different shot types 
summary(as.factor(chelseaShotsData$shot.type.name))

# Look at what period shots are taken 
summary(as.factor(chelseaShotsData$period))
```

```{r}
table(chelseaShotsData$shot.type.name, chelseaShotsData$period)
```

```{r}
create_Pitch(grass_colour = "#538032", 
             line_colour =  "#ffffff", 
             background_colour = "#538032", 
             goal_colour = "#000000") + 
  geom_point(data = chelseaShotsData, aes(x = location.x , # Set x-axis as x coordinate
                 y = location.y,  # Set y-axis as y coordinate
                 color = factor(shot_res)), # Set color as shot result
                 alpha = 0.3,
             size = 3) + # Set point size
   scale_color_manual(values = c("1" = "blue", "0" = "red"), # Scale colors manually
                     labels = c("1" = "Goal", "0" = "Miss")) +  # Set color labels
  guides(alpha = FALSE) + # Turn off alpha legend 
  labs(color = "Shot Result:", title = "Chelsea FCW Shots", subtitle = "WSL, 2019-20") + 
  scale_y_reverse() + 
  coord_fixed(ratio = 105/100)
```

```{r}
# Create x and y grid vectors
x_vals <- c(55,60,65,70,75, 80,85, 90, 95, 100, 105, 110, 115)
y_vals <- c(10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65)
# Use expand.grid to create data frame with each combination of different values
grid_vals <- expand.grid(x_vals, y_vals)
# Change names for grid
names(grid_vals) <- c("xmin", "ymin")
# Create maximum grid value columns
grid_vals$xmax <- grid_vals$xmin + 5
grid_vals$ymax <- grid_vals$ymin + 5
# View first five rows of grid
head(grid_vals)
```

```{r}
# Create empty vectors to store shots and goals
shots <- goals <- rep(NA, # Repeat NA 
                      times = nrow(grid_vals)) # The number of rows there are in grid values
# for each row of the grid values
for(i in 1:nrow(grid_vals)){
  # Calculate number of shots taken inside each grid box
  shots[i] <- sum(chelseaShotsData$location.y >= grid_vals$ymin[i] & # y greater than or equal to minimum y
                 chelseaShotsData$location.y < grid_vals$ymax[i] & # y less than maximum y
                 chelseaShotsData$location.x >= grid_vals$xmin[i] & # x greater than or equal to minimum y
                 chelseaShotsData$location.x < grid_vals$xmax[i]) # x less than maximum y
  # Calculate number of goals scored inside each grid box
 goals[i] <- sum(chelseaShotsData$location.y >= grid_vals$ymin[i] & # y greater than or equal to minimum y
                 chelseaShotsData$location.y < grid_vals$ymax[i] & # y less than maximum y
                 chelseaShotsData$location.x >= grid_vals$xmin[i] & # x greater than or equal to minimum y
                 chelseaShotsData$location.x < grid_vals$xmax[i] & # x less than maximum y
                 chelseaShotsData$shot_res == 1) # Shot result equal to 1
}
```

```{r}
# Calculate rate of scoring from each grid box
rate <- goals/shots
# Add columns to data frame
grid_vals$shots <- shots
grid_vals$goals <- goals
grid_vals$rate <- rate
# Select only grid sections where at least one shot is attempted.
grid_vals_use <- grid_vals[grid_vals$shots >= 1,]
# Print out grid values
grid_vals_use
```

```{r}
create_Pitch(grass_colour = "#538032", 
             line_colour =  "#ffffff", 
             background_colour = "#538032", 
             goal_colour = "#000000") + 
  geom_rect(data = grid_vals_use, aes(xmin = xmin,  # Set minimum x value
                xmax = xmax ,  # Set maximum x value
                ymin =  ymin, # Set minimum y value
                ymax = ymax,  # Set maximum y value
                fill = rate), # Set fill of each rectangle
              alpha = 0.9) + # Set transperancy
  scale_fill_gradient(low = "blue", high = "red") + # Scale fill gradien
  labs(fill = "Scoring Rate:", title = "Chelsea FCW Scoring Rate by Pitch Area", subtitle = "WSL, 2019-20") + 
  scale_y_reverse() + 
  coord_fixed(ratio = 105/100)
```

```{r}
# Extract unique players from dataset
players <- unique(chelseaShotsData$player.name)
# Create empty vectors to store values
shots <- goals <- xg_vals <- rep(NA, length(players))
# For each player
for(i in 1:length(players)){
  # Sum up number of shots the player had
  shots[i] <- sum(chelseaShotsData$player.name == players[i])
  # Sum up number of goals the player scored
  goals[i] <- sum(chelseaShotsData$player.name == players[i] &
                    chelseaShotsData$shot_res == 1)
  # Sum up XG for the players shots
  xg_vals[i] <- sum(chelseaShotsData$shot.statsbomb_xg[chelseaShotsData$player.name == players[i]])
}
# Calculate difference in players expected vs actual goals
xg_goal_diff <- xg_vals - goals
# Join results together
player_xg_db <- cbind.data.frame(players, shots, goals, xg_vals, xg_goal_diff)
```

```{r}
# Show top of dataframe
head(player_xg_db[order(player_xg_db$goals, decreasing = TRUE),], # Order rows by goals scored
     n=10) # Set n as 10 to show top 10 players
```

Lets view the players who out performed in terms of expected goals. We can do this by ordering the data using the difference between actual and expected goals.

```{r}
# View top of data frame
head(player_xg_db[order(player_xg_db$xg_goal_diff),], # Order by expected and actual goal difference
     n=22) # Set n as 10 to show 10 rows
```

And players who under performed, here we set decreasing as TRUE to show highest values first:

```{r}
# View top of data frame
head(player_xg_db[order(player_xg_db$xg_goal_diff, # Order by actual v expected goal difference
                        decreasing = TRUE),], # Decreasing values so highest values first
     n=10) # Set n as 10 to show top 10 rows
```

## Shot Context -- 5 rows/events before a shot

```{r}
# Function to get 5 rows before and including "Shot"
shotContext <- function(data) {
  filtered_data <- data |> 
    filter(team.name == "Chelsea FCW") # Filter for Chelsea FCW events
  
  if (nrow(filtered_data) == 0) {
    warning("No data found for Chelsea FCW.")
    return(data.frame()) # Return an empty data frame
  }
  
  shot_rows <- filtered_data |> 
    mutate(row_number = row_number()) |> # Add a row number
    filter(type.name == "Shot") |>       # Identify rows with "Shot"
    pull(row_number)                     # Extract their row numbers
  
  if (length(shot_rows) == 0) {
    warning("No 'Shot' events found for Chelsea FCW.")
    return(data.frame()) # Return an empty data frame
  }
  
  # Get rows within the 5 rows before each "Shot"
  row_indices <- shot_rows |> 
    map(~seq(max(. - 5, 1), .)) |>       # Generate sequences for each "Shot"
    unlist() |>                          # Flatten the list into a single vector
    unique() |>                          # Remove duplicate row indices
    sort()                               # Sort the row indices
  
  # Subset the original data using the calculated row indices
  filtered_data[row_indices, ]
}

beforeShots <- shotContext(chelseaAttackData)

print(beforeShots)
```

## Gini Index by Position -- Do Chelsea players tend to carry(turn, move, etc) or pass under pressure? 

### For under pressure

```{r}
# Create function that collects the action frequency dependfing on action type  
actionFreqUnderPressure <- function(data, player_name, action_type) {
  actionFreq <- 0
  
  for (i in 1:nrow(data)) {
    # Check if the conditions are met (ignoring missing values with isTRUE())
    if (isTRUE(data$player.name[i] == player_name) && 
        isTRUE(data$under_pressure[i] == TRUE) && 
        isTRUE(data$type.name[i] == action_type)) {
      actionFreq <- actionFreq + 1
    }
  }
  return(actionFreq)
}
```

```{r}
# Calculate pass frequency under pressure for each player
passResult1 <- sapply(chelseaNames, function(player) {
  actionFreqUnderPressure(chelseaAttackData, player, "Pass")
})

# Calculate carry frequency under pressure for each player
carryResult1 <- sapply(chelseaNames, function(player) {
  actionFreqUnderPressure(chelseaAttackData, player, "Carry")
})

freqResult1 <- data.frame(
  player = chelseaNames,
  passFrequencyPressure = passResult1,
  carryFrequencyPressure = carryResult1
)

print(freqResult1)
```

```{r}
# Group players by position (defender, midfielder, forward)
unique(chelseaAttackData$position.name)

defensePos <- c("Left Center Back", "Left Back", "Right Back", "Right Center Back", "Right Wing Back", "Center Back", "Left Wing Back")
midfieldPos <- c("Right Center Midfield", "Left Midfield", "Right Midfield", "Left Center Midfield", 
                 "Center Defensive Midfield", "Left Wing", "Right Wing", "Center Attacking Midfield", 
                 "Left Defensive Midfield", "Right Defensive Midfield")
forwardPos <- c("Left Center Forward", "Right Center Forward", "Center Forward")

```

```{r}
# Join frequencies with position data and classify by position category
freqResult1 <- merge(freqResult1, unique(chelseaAttackData[, c("player.name", "position.name")]), by.x = "player", 
                     by.y = "player.name", all.x = TRUE)

# Add a new column for position category
freqResult1$positionCategory <- ifelse(freqResult1$position.name %in% defensePos, "Defender",
                               ifelse(freqResult1$position.name %in% midfieldPos, "Midfielder",
                               ifelse(freqResult1$position.name %in% forwardPos, "Forward", NA)))

# Exclude goalkeepers from the frequency results
freqResult1 <- freqResult1[freqResult1$position.name != "Goalkeeper", ]
```

```{r}
# Remove duplicate rows based on player name and position
freqResult1 <- freqResult1 |>
  distinct(player, positionCategory, .keep_all = TRUE)

print(freqResult1)
```

```{r}
# Count frequency of each player-position combination, and select the max for each player
positionFrequency <- chelseaAttackData |>
  filter(team.name == "Chelsea FCW") |>    
  group_by(player.name, position.name) |>
  summarise(position_count = n(), .groups = 'drop') |>
  group_by(player.name) |>
  slice_max(position_count, n = 1) |>      # Keep only the row with the highest count for each player
  ungroup()

print(positionFrequency)
```

```{r}
positionFrequency$positionCategory <- ifelse(positionFrequency$position.name %in% defensePos, "Defender",
                                      ifelse(positionFrequency$position.name %in% midfieldPos, "Midfielder",
                                      ifelse(positionFrequency$position.name %in% forwardPos, "Forward", NA)))

# Exclude goalkeepers
positionFrequency <- positionFrequency[positionFrequency$position.name != "Goalkeeper", ]

# Merge freqResult1 with positionFrequency
freqResult1 <- merge(freqResult1, 
                     positionFrequency[, c("player.name", "positionCategory")], 
                     by.x = "player", by.y = "player.name", all.x = TRUE)

# Check if columns exist before filtering or renaming
if ("positionCategory.x" %in% colnames(freqResult1) && "positionCategory.y" %in% colnames(freqResult1)) {
    # Filter rows where positionCategory.x is equal to positionCategory.y
    freqResult1 <- freqResult1[freqResult1$positionCategory.x == freqResult1$positionCategory.y, ]

    # Remove unnecessary columns and rename properly
    freqResult1 <- freqResult1 |>
        select(-positionCategory.x) |>
        rename(positionCategory = positionCategory.y)
}

print(freqResult1)
```

### For not under pressure

```{r}
# Create function that collects the action frequency dependfing on action type
actionFreqNotUnderPressure <- function(data, player_name, action_type) {
  actionFreq <- 0
  
  for (i in 1:nrow(data)) {
    # Check if the conditions are met
    if (isTRUE(data$player.name[i] == player_name) &&
        is.na(data$under_pressure[i]) &&
        isTRUE(data$type.name[i] == action_type)) {
      actionFreq <- actionFreq + 1
    }
  }
  return(actionFreq)
}
```

```{r}
# Calculate pass frequency under pressure for each player
passResult2 <- sapply(chelseaNames, function(player) {
  actionFreqNotUnderPressure(chelseaAttackData, player, "Pass")
})

# Calculate carry frequency under pressure for each player
carryResult2 <- sapply(chelseaNames, function(player) {
  actionFreqNotUnderPressure(chelseaAttackData, player, "Carry")
})

freqResult2 <- data.frame(
  player = chelseaNames,
  passFrequencyNoPressure = passResult2,
  carryFrequencyNoPressure = carryResult2
)

print(freqResult2)
```

```{r}
# Merge freqResult1 and freqResult2 by the 'player' column
mergedResults <- merge(freqResult1, freqResult2, by = "player")

mergedResults <- mergedResults |>
  select(-position.name)

print(mergedResults)
```


### Adding under pressure and not under pressure to same data frame

```{r}
# Calculate Gini indices for each position group
giniResults <- mergedResults |>
  group_by(positionCategory) |>
  summarise(
    gini_passPressure = Gini(passFrequencyPressure),
    gini_carryPressure = Gini(carryFrequencyPressure),
    gini_passNoPressure = Gini(passFrequencyNoPressure),
    gini_carryNoPressure = Gini(carryFrequencyNoPressure)
  )

print(giniResults)
```
```{r}
# Reshape data for ggplot2
giniMelt <- melt(giniResults, id.vars = "positionCategory",
                 variable.name = "ActionContext", value.name = "GiniIndex")

# Separate Action and Pressure Context
giniMelt$Action <- ifelse(grepl("pass", giniMelt$ActionContext), "Pass", "Carry")
giniMelt$Pressure <- ifelse(grepl("UnderPressure", giniMelt$ActionContext), "Under Pressure", "No Pressure")
```

```{r}
ggplot(giniMelt, aes(x = Pressure, y = Action)) +
  geom_tile(aes(fill = GiniIndex)) +
  facet_wrap(~ positionCategory) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = median(giniMelt$GiniIndex, na.rm = TRUE)
  ) +
  labs(
    title = "Chelsea Gini Index for Actions Under/No Pressure by Positional Group",
    x = "Pressure Context",
    y = "Action Type",
    fill = "Gini Index"
  ) +
  theme_minimal()
```

## Rest of League Gini Index by Position Group 

### Finding team names in data set

```{r}
unique(statsBombData$team.name)
```

### Creating new dataset 

```{r}
otherData <- statsBombData |>
  filter(team.name != "Chelsea FCW") 

otherNames <- unique(statsBombData$player.name)
otherNames <- na.omit(otherNames)
```

```{r}
# Calculate pass frequency under pressure for each player
passResult5 <- sapply(otherNames, function(player) {
  actionFreqUnderPressure(otherData, player, "Pass")
})

# Calculate carry frequency under pressure for each player
carryResult5 <- sapply(otherNames, function(player) {
  actionFreqUnderPressure(otherData, player, "Carry")
})

freqResult5 <- data.frame(
  player = otherNames,
  passFrequencyPressure = passResult5,
  carryFrequencyPressure = carryResult5
)

head(freqResult5)
```

```{r}
# Group players by position (defender, midfielder, forward)
unique(otherData$position.name)

defensePos <- c("Left Center Back", "Left Back", "Right Back", "Right Center Back", "Right Wing Back", "Center Back", "Left Wing Back")
midfieldPos <- c("Right Center Midfield", "Left Midfield", "Right Midfield", "Left Center Midfield", 
                 "Center Defensive Midfield", "Left Wing", "Right Wing", "Center Attacking Midfield", 
                 "Left Defensive Midfield", "Right Defensive Midfield", "Right Attacking Midfield", "Left Attacking Midfield")
forwardPos <- c("Left Center Forward", "Right Center Forward", "Center Forward")

```

```{r}
# Join frequencies with position data and classify by position category
freqResult5 <- merge(freqResult5, unique(otherData[, c("player.name", "position.name")]), by.x = "player", 
                     by.y = "player.name", all.x = TRUE)

# Add a new column for position category
freqResult5$positionCategory <- ifelse(freqResult5$position.name %in% defensePos, "Defender",
                               ifelse(freqResult5$position.name %in% midfieldPos, "Midfielder",
                               ifelse(freqResult5$position.name %in% forwardPos, "Forward", NA)))

# Exclude goalkeepers and substitutes from the frequency results
freqResult5 <- freqResult5[freqResult5$position.name != "Goalkeeper" & freqResult5$position.name != "Substitute", ]
```

```{r}
# Remove duplicate rows based on player name and position
freqResult5 <- freqResult5 |>
  distinct(player, positionCategory, .keep_all = TRUE)

print(freqResult5)
```

```{r}
# Count frequency of each player-position combination, and select the max for each player
positionFrequency2 <- otherData |>
  group_by(player.name, position.name) |>
  summarise(position_count = n(), .groups = 'drop') |>
  group_by(player.name) |>
  slice_max(position_count, n = 1) |>      # Keep only the row with the highest count for each player
  ungroup()

print(positionFrequency2)
```

```{r}
positionFrequency2$positionCategory <- ifelse(positionFrequency2$position.name %in% defensePos, "Defender",
                                      ifelse(positionFrequency2$position.name %in% midfieldPos, "Midfielder",
                                      ifelse(positionFrequency2$position.name %in% forwardPos, "Forward", NA)))

# Exclude goalkeepers
positionFrequency2 <- positionFrequency2[positionFrequency2$position.name != "Goalkeeper", ]

# Merge freqResult1 with positionFrequency
freqResult5 <- merge(freqResult5, 
                     positionFrequency2[, c("player.name", "positionCategory")], 
                     by.x = "player", by.y = "player.name", all.x = TRUE)

# Check if columns exist before filtering or renaming
if ("positionCategory.x" %in% colnames(freqResult5) && "positionCategory.y" %in% colnames(freqResult5)) {
    # Filter rows where positionCategory.x is equal to positionCategory.y
    freqResult5 <- freqResult5[freqResult5$positionCategory.x == freqResult5$positionCategory.y, ]

    # Remove unnecessary columns and rename properly
    freqResult5 <- freqResult5 |>
        select(-positionCategory.x) |>
        rename(positionCategory = positionCategory.y)
}

print(freqResult5)
```

```{r}
# Calculate pass frequency under pressure for each player
passResult6 <- sapply(otherNames, function(player) {
  actionFreqNotUnderPressure(otherData, player, "Pass")
})

# Calculate carry frequency under pressure for each player
carryResult6 <- sapply(otherNames, function(player) {
  actionFreqNotUnderPressure(otherData, player, "Carry")
})

freqResult6 <- data.frame(
  player = otherNames,
  passFrequencyNoPressure = passResult6,
  carryFrequencyNoPressure = carryResult6
)

head(freqResult6)
```

```{r}
# Merge freqResult1 and freqResult2 by the 'player' column
mergedResults2 <- merge(freqResult5, freqResult6, by = "player")

mergedResults2 <- mergedResults2 |>
  select(-position.name)

print(mergedResults2)
```

```{r}
# Calculate Gini indices for each position group
giniResults3 <- mergedResults2 |>
  group_by(positionCategory) |>
  summarise(
    gini_passPressure = Gini(passFrequencyPressure),
    gini_carryPressure = Gini(carryFrequencyPressure),
    gini_passNoPressure = Gini(passFrequencyNoPressure),
    gini_carryNoPressure = Gini(carryFrequencyNoPressure)
  )

print(giniResults3)
```

```{r}
# Reshape data for ggplot2
giniMelt3 <- melt(giniResults3, id.vars = "positionCategory",
                 variable.name = "ActionContext", value.name = "GiniIndex")

# Separate Action and Pressure Context
giniMelt3$Action <- ifelse(grepl("pass", giniMelt$ActionContext), "Pass", "Carry")
giniMelt3$Pressure <- ifelse(grepl("UnderPressure", giniMelt$ActionContext), "Under Pressure", "No Pressure")
```

```{r}
ggplot(giniMelt3, aes(x = Pressure, y = Action)) +
  geom_tile(aes(fill = GiniIndex)) +
  facet_wrap(~ positionCategory) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = median(giniMelt$GiniIndex, na.rm = TRUE)
  ) +
  labs(
    title = "League Average Gini Index for Actions Under/No Pressure by Positional Group",
    x = "Pressure Context",
    y = "Action Type",
    fill = "Gini Index"
  ) +
  theme_minimal()
```

## Gini Expansion by Player 

```{r}
combineActionFrequencies <- function(data, action_types) {
  results <- list()

  # Loop through each action type
  for (action_type in action_types) {
    # Calculate frequencies for under pressure
    underPressure <- data |>
      filter(under_pressure == TRUE, type.name == action_type) |>
      group_by(match_id, player.name) |>
      summarise(
        !!paste0(action_type, "FrequencyUnderPressure") := n(),
        .groups = "drop"
      )
    
    # Calculate frequencies for not under pressure
    notUnderPressure <- data |>
      filter(is.na(data$under_pressure), type.name == action_type) |>
      group_by(match_id, player.name) |>
      summarise(
        !!paste0(action_type, "FrequencyNotUnderPressure") := n(),
        .groups = "drop"
      )
    
    # Merge under pressure and not under pressure results
    combined <- full_join(underPressure, notUnderPressure, by = c("match_id", "player.name"))
    
    # Append to results list
    results[[action_type]] <- combined
  }
  
  # Merge results for all action types into a single data frame
  finalResult <- reduce(results, full_join, by = c("match_id", "player.name"))
  
  # Replace NAs with 0 (optional)
  finalResult[is.na(finalResult)] <- 0
  
  return(finalResult)
}
```

```{r}
actionTypes <- c("Pass", "Carry")
combinedPlayerResults <- combineActionFrequencies(chelseaPlayerData, actionTypes)
print(combinedPlayerResults)
```

```{r}
# Create function for calculating Gini index 
calculateGiniIndices <- function(data, player_col, pass_pressure_col, carry_pressure_col, pass_no_pressure_col, carry_no_pressure_col) {
  gini_results <- data |>
    group_by({{player_col}}) |>
    summarise(
      gini_passPressure = Gini({{pass_pressure_col}}),
      gini_carryPressure = Gini({{carry_pressure_col}}),
      gini_passNoPressure = Gini({{pass_no_pressure_col}}),
      gini_carryNoPressure = Gini({{carry_no_pressure_col}})
    )
  return(gini_results)
}
```

```{r}
giniResults2 <- calculateGiniIndices(data = combinedPlayerResults, 
                     player_col = player.name,
                     pass_pressure_col = PassFrequencyUnderPressure,
                     carry_pressure_col = CarryFrequencyUnderPressure,
                     pass_no_pressure_col = PassFrequencyNotUnderPressure,
                     carry_no_pressure_col = CarryFrequencyNotUnderPressure)

print(giniResults2)
```

```{r}
# Reshape data
giniMelt2 <- melt(giniResults2, id.vars = "player.name", variable.name = "ActionContext", value.name = "GiniIndex")

# Correctly label the Pressure context based on the ActionContext
giniMelt2$Action <- ifelse(grepl("pass", giniMelt2$ActionContext), "Pass", "Carry")
giniMelt2$Pressure <- ifelse(grepl("Pressure", giniMelt2$ActionContext), "Under Pressure", "No Pressure")
giniMelt2$Pressure <- ifelse(grepl("NoPressure", giniMelt2$ActionContext), "No Pressure", "Under Pressure")
```


```{r}
ggplot(giniMelt2, aes(x = Pressure, y = Action)) +
  geom_tile(aes(fill = GiniIndex)) +
  facet_wrap(~ player.name, ncol = 4) + # Increase columns for less overlap
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = median(giniMelt2$GiniIndex, na.rm = TRUE)
  ) +
  labs(
    title = "Gini Index for Actions Under/No Pressure by Player",
    x = "Pressure Context",
    y = "Action Type",
    fill = "Gini Index"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6), # Rotate x-axis for clarity
    strip.text = element_text(size = 5), # Adjust facet text size
    axis.text.y = element_text(size = 6)
  ) + 
  scale_x_discrete(
    limits = c("Under Pressure", "No Pressure")) 

```



